Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants \thanks{Supported by EPSRC grant EP/J012564/1, ERC project 280053 (CPROVER) and the H2020 FET OPEN SC$^2$.}

### 2017-02-11 hscc2017/paper.tex
hscc17

%------------------------------------
\subsection{Verifying Closed-loop Control Systems}
\label{verifying-closed-loop-control-systems}
%------------------------------------

@@

  \path[->] (v1) edge (q1.north);
  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (q2.west)
   (q2.east) edge (zoh.west)
   (zoh.east) edge (gs.west)
   (switch2) edge (q1.west)
   (q1.east) edge (aux1.west)
   (gs.east) edge (switch1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Closed-loop digital control system\label{fig:sampledsystem}}
\end{figure*}

@@

A typical digital control system with a continuous plant and a discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantization errors (notice that each of them
may have a different FWL than the controller),
%\red{[why is the DAC introducing a quantisation error again?]}
which are modeled as noises
$\nu_{1}(z)$ and $\nu_{2}(z)$, $G(s)$ is the continuous-time plant model
with parametric additive uncertainty $\Delta_p{G}_n(s)$ and
$\Delta_p{G}_d(s)$, $R(z)$ is the reference signal, $U(z)$ is the control
signal, and $\hat{Y}(z)$ is the output signal affected by the noises and
uncertainties in the closed-loop system.
%
%
%
The ADC and DAC may be abstracted by transforming the closed-loop system in
Figure~\ref{fig:sampledsystem} into the digital system in 
Figure~\ref{fig:digsystem1}, where the parametric uncertainties are
represented by~$\Delta_p \vec{G}$
and the effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
seen as additive noise.

@@

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu2) edge (circle2.north);
  \path[->] (nu1) edge (circle3.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (circle2.west)
   (circle2.east) edge (gs.west)
   (gs.east) edge (circle3.west)
   (circle3.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
\caption{Digital equivalent to the system in Figure~\ref{fig:sampledsystem} 
\label{fig:digsystem1}}
\end{figure}

%\red{[why is the output in Figure 2 tilde, whereas that in  figure 1 hat?]}
%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors, and uncertainties
introduced by quantizations in the ADC and
DAC conversions.  These sources are assumed to be non-deterministic, 
which poses no further restriction to our model, thus, the uncertain model
may be rewritten as a vector of coefficients in the z-domain using equation
(\ref{eq:hatgvector})
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties. Notice that the quantization noise is additive, which means
it does not affect the transfer function. Instead, we prove that the system
is stable given these non-deterministic noise inputs.

@@

Thus, the perturbed controller model $\hat{C}(z)$ may be obtained from the
original model $C(z)=\frac{C_{n}(z)}{C_{d}(z)}$ as follows:
%
\begin{equation}
\hat{C}(z)=\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))=\frac{\mathcal{F}_{M_C}\langle I,F \rangle(C_n(z))}{\mathcal{F}_{N_C}\langle I,F \rangle(C_d(z))}.
\end{equation}
%
In the case of a synthesized controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.

@@

%------------------------------------
\subsection{Stability under Finite Representation and Quantisation Noise}
\label{sec:stability}
%------------------------------------

Correct synthesis of the digital controller requires consideration of the
effect of FWL on the controller and quantization noises in the closed loop
system.  Let the quantizer $Q1$ be the source of a white noise $\nu_{1}$ and
$Q2$ be the source of a white noise $\nu_{2}$.  The following equation
models the system in Figure~\ref{fig:sampledsystem}, encompassing the
parametric uncertainties $\Delta \vec{G}$ and the FWL effects on
the controller $\hat{C}(z)$:
%
\begin{equation}
\hat{Y}(z)=\nu_{1}(z)+\hat{G}(z)\cdot C(z)\cdot R(z)+\hat{G}(z)\cdot\nu_{2}(z)-\hat{G}(z)\cdot \hat{C}(z)\cdot \hat{Y}(z)
\end{equation}
%
%
%
%
The above can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}(z)\nu_{1}(z)+H_{2}(z)\cdot\nu_{2}(z)+H_{3}(z)\cdot R(z),
\end{equation}
%
where
%
\begin{equation*}
H_{1}(z)=\frac{1}{1+\hat{G}(z)\cdot \hat{C}(z)}
\end{equation*}
%
\begin{equation*}
H_{2}(z)=\frac{\hat{G}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}
\end{equation*}
%
and
%
\begin{equation*}
H_{3}(z)=\frac{\hat{G}(z)\cdot \hat{C}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}
\end{equation*}

\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from $Q1$) and $\nu_{2}$ (from $Q2$) are
uncorrelated white noises and their amplitudes are always bounded by the
half of quantization step~\cite{astrom1997computer}, i.e., $\vert \nu_{1}
\vert \leq \frac{q_{1}}{2}$ and $\vert \nu_{2} \vert \leq \frac{q_{2}}{2}$,
where $q_{1}$ and $q_{2}$ are the quantization steps of ADC and DAC, respectively.
% 
\end{myassumption}

@@

A discrete-time linear system is said to be Bounded-Input and Bounded-Output
(BIBO) stable if and only if every pole of its transfer function lies inside
the unit circle \cite{Astrom08}.  Analyzing Eq.~\eqref{eq:outputfunctions},
the following proposition gives necessary conditions for the closed-loop system 
BIBO stability in Figure~\ref{fig:sampledsystem}, with regards to the 
exogenous signals $R(z)$, $\nu_{1}$, and $\nu_{2}$, which are bounded.

\begin{myprop}{\cite{Bessa16,fadali}}
\label{prop:eq_int_stab} 
%
Consider a feedback closed-loop control system as given in
Figure~\ref{fig:sampledsystem} with a FWL implementation of the digital
controller $\hat{C}(z) =\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))$ and uncertain discrete
model of the plant
%
$$
\hat{G}(z) =\frac{\hat{G}_n(z)}{\hat{G}_d(z)}: \vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}
\text{\quad(from~\ref{eq:coefficients},\ref{eq:delta_coefficients})}.
$$
%
Then $\hat{G}(z)$ is BIBO-stable if and only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=\mathcal{F}_{M_C}\langle I,F \rangle(C_n(z))\hat{G}_n(z)+\mathcal{F}_{N_C}\langle I,F \rangle(C_d(z))\hat{G}_d(z);
\end{equation}
\item the direct loop product $\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))\cdot \hat{G}(z)$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{myprop}

Proposition~\ref{prop:eq_int_stab} provides necessary (and sufficient) conditions for the
controller to stabilize the closed-loop system, considering plant
parametric uncertainties (i.e., $\Delta_p \vec{G}$), quantization noises
($\nu_{1}$ and $\nu_{2}$) and FWL effects in the control software.  Note
that the model for quantization noise enters as a signal to be stabilized (i.e.,
the noise may be disregarded if the conditions on
Proposition~\ref{prop:eq_int_stab} are satisfied and if the quantization
noise is bounded.  The bound on the quantization noise is given by
Assumption~\ref{whitenoise}.

>>>>

%------------------------------------
\subsection{Verifying Closed-loop Control Systems \\ \blue{Effects of Imprecisions, Quantization, and Finite Word Length Representation}}
\label{verifying-closed-loop-control-systems}
%------------------------------------

@@

  \path[->] (v1) edge (q1.north);
  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (q2.west)
   (q2.east) edge (zoh.west)
   (zoh.east) edge (gs.west)
   (switch2) edge (q1.west)
   (q1.east) edge (aux1.west)
   (gs.east) edge (switch1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Closed-loop digital control system \red{[has this notation for the plant ever been introduced?]} \label{fig:sampledsystem}}
\end{figure*}

@@

A typical digital control system with a continuous plant and a discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantization errors (notice that each of them
may have a different FWL representation than the controller),
%\red{[why is the DAC introducing a quantisation error again?]}
which are modeled as noises
$\nu_{1}(z)$ and $\nu_{2}(z)$;  
$G(s)$ is the continuous-time plant model
with parametric additive uncertainty $\Delta_p{G}_n(s)$ and $\Delta_p{G}_d(s)$ (as mentioned above);  
$R(z)$ is a given reference signal; 
$U(z)$ is the control signal; 
and $\hat{Y}(z)$ is the output signal affected by the noises and uncertainties in the closed-loop system 
\red{[notice we have tilde outputs in the two figures! these need to be explained here.]}. 
%
%
%
The ADC and DAC may be abstracted by transforming the closed-loop system in
Figure~\ref{fig:sampledsystem} into the digital system in 
Figure~\ref{fig:digsystem1}, 
where 
% the parametric uncertainties are represented by~$\Delta_p \vec{G}$ and 
the effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
seen as additive noise.

@@

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu2) edge (circle2.north);
  \path[->] (nu1) edge (circle3.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (circle2.west)
   (circle2.east) edge (gs.west)
   (gs.east) edge (circle3.west)
   (circle3.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
\caption{Fully digital equivalent to system in Figure~\ref{fig:sampledsystem} 
\label{fig:digsystem1}}
\end{figure}

%\red{[why is the output in Figure 2 tilde, whereas that in  figure 1 hat?]}
%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors  
(which are represented by~$\Delta_p \vec{G}$),  
and uncertainties introduced by quantizations in the ADC and
DAC conversions \red{[how are these denoted?]}.  
These sources are assumed to be non-deterministic \blue{[here you mean specifically the quantization effects?]}:  
%which poses no further restriction to our model, 
thus, 
the uncertain model may be rewritten as a vector of coefficients in the z-domain using equation (\ref{eq:hatgvector}) as 
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}. 
\end{equation}
%
%where $\Delta_p \vec{G}$ is the term that corresponds to the parametric uncertainties. 
Recall that we discussed how the quantization noise is encompassed as an additive term, 
which means it does not enter parametrically in the transfer function. 
Instead, we later show that the system is stable given these non-deterministic noise inputs.

@@

Thus, the perturbed controller model $\hat{C}(z)$ may be obtained from the
original model $C(z)=\frac{C_{n}(z)}{C_{d}(z)}$ as follows:
%
\begin{equation}
\hat{C}(z)=\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))=\frac{\mathcal{F}_{M_C}\langle I,F \rangle(C_n(z))}{\mathcal{F}_{N_C}\langle I,F \rangle(C_d(z))}.
\end{equation}
%
\blue{[not confusing next?]}
In the case of a synthesized controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.

@@

%------------------------------------
\subsection{Stability under FWL Representation and under Quantisation Noise \blue{(and params uncertainties?)}}
\label{sec:stability}
%------------------------------------

Sound synthesis of the digital controller requires the consideration of the effect of FWL on the controller, 
and of quantization noises in the closed-loop system.  
Let the quantizer $Q1$ \blue{[new notation- needed really?] (DAC)} be the source of a white noise $\nu_{1}$, 
and $Q2$ \blue{[ditto] (ADC)} be the source of a white noise $\nu_{2}$.  The following equation
models the system in Figure~\ref{fig:sampledsystem}, encompassing the
parametric uncertainties $\Delta \vec{G}$ and the FWL effects on
the controller $\hat{C}(z)$:
%
\begin{equation}
\hat{Y}(z)=\nu_{1}(z)+\hat{G}(z)\cdot C(z)\cdot R(z)+\hat{G}(z)\cdot\nu_{2}(z)-\hat{G}(z)\cdot \hat{C}(z)\cdot \hat{Y}(z). 
\end{equation}
%
The above can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}(z)\cdot\nu_{1}(z)+H_{2}(z)\cdot\nu_{2}(z)+H_{3}(z)\cdot R(z),
\end{equation}
%
where
%
\begin{equation*}
H_{1}(z)=\frac{1}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{2}(z)=\frac{\hat{G}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{3}(z)=\frac{\hat{G}(z)\cdot \hat{C}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}. 
\end{equation*}

\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ \blue{(from $Q1$)} and $\nu_{2}$ \blue{(from $Q2$)} are
uncorrelated white noises and their amplitudes are always bounded by the
half of quantization step~\cite{astrom1997computer}, i.e., $\vert \nu_{1}
\vert \leq \frac{q_{1}}{2}$ and $\vert \nu_{2} \vert \leq \frac{q_{2}}{2}$,
where $q_{1}$ and $q_{2}$ are the quantization steps of ADC and DAC, respectively.
% 
\end{myassumption}

@@

A discrete-time dynamical system is said to be Bounded-Input and Bounded-Output
(BIBO) stable if bounded inputs necessarily result in bounded outputs. 
This condition holds true over an LTI model if and only if every pole of its transfer function lies inside the unit circle \cite{Astrom08}.  
Analyzing Eq.~\eqref{eq:outputfunctions}, 
the following proposition provides conditions for the BIBO stability of the system in Figure~\ref{fig:sampledsystem}, 
with regards to the exogenous signals $R(z)$, $\nu_{1}$, and $\nu_{2}$, which are all bounded 
(in particular, the bound on the quantization noise is given by
Assumption~\ref{whitenoise}). 
%
\begin{myprop}{\cite{Bessa16,fadali}}
\label{prop:eq_int_stab} 
%
Consider a feedback closed-loop control system as given in
Figure~\ref{fig:sampledsystem} with a FWL implementation of the digital
controller $\hat{C}(z) =\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))$ and uncertain discrete
model of the plant from~\eqref{eq:coefficients},\eqref{eq:delta_coefficients} 
%
$$
\hat{G}(z) =\frac{\hat{G}_n(z)}{\hat{G}_d(z)}, \quad \vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}.
$$
%
Then $\hat{G}(z)$ is BIBO-stable if and only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
\hspace{-3ex} S(z)=\mathcal{F}_{M_C}\langle I,F \rangle(C_n(z))\hat{G}_n(z)+\mathcal{F}_{N_C}\langle I,F \rangle(C_d(z))\hat{G}_d(z);
\end{equation}
\item the direct loop product $\mathcal{F}_{M_C,N_C}\langle I,F \rangle(C(z))\cdot \hat{G}(z)$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{myprop}

Proposition~\ref{prop:eq_int_stab} provides necessary (and sufficient) conditions for the
controller to stabilize the closed-loop system, considering plant
parametric uncertainties (i.e., $\Delta_p \vec{G}$), quantization noises
($\nu_{1}$ and $\nu_{2}$) and FWL effects in the control software.  
In particular, 
note that the model for quantization noise enters as a signal to be stabilized: 
in practice, 
if the quantization noise is bounded, 
the noise may be disregarded if the conditions on
Proposition~\ref{prop:eq_int_stab} are satisfied.
